---
title: "Carceral Facilities and Their OSHA Hazard and Toxicity Record"
author: "Savannah Hunter"
date: "6/3/2021"
output:
  html_document: default
  pdf_document: default
---

# Load Packages and Data

```{r}
#### Load in Packages ####
xfun::session_info('DT')
library(DT)
library("tidyverse")
library("lubridate")

```

```{r}

#### Load in Data ####
caprisons = read.csv(file = "CA_prison_insp_viol_2010_op_fac_long.csv", header = TRUE, stringsAsFactors = FALSE)

```

# Calculate the Statistics in the Long Dataset

## Penalties

```{r}
# Total Initial Penalty for Establishments

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$iptotal = sum(estb$initial_penalty, na.rm = TRUE)
  return(estb)
}))


# Total Current Penalty for Establishments 

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$cptotal = sum(estb$current_penalty, na.rm = TRUE)
  return(estb)
}))

# Average Initial Penalty for Establishments 
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$ipavg = mean(estb$initial_penalty, na.rm = TRUE)
  return(estb)
}))
    
  # NAN appears for rows where the facility had no initial penalty and is thus     dividing by zero. I convert these to zero. 
  caprisons$ipavg = as.numeric(gsub("NaN", 0, caprisons$ipavg))

  # format to two decimals
  caprisons$ipavg = as.numeric(formatC(caprisons$ipavg, digits = 2, format = "f"))
  
  
# Average Current Penalty for Establishments
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$cpavg = mean(estb$current_penalty, na.rm = TRUE)
  return(estb)
}))

  # NAN appears for rows where the facility had no initial penalty and is thus     dividing by zero. I convert these to zero. 
  caprisons$cpavg = as.numeric(gsub("NaN", 0, caprisons$cpavg))
  
    # format to two decimals
   caprisons$cpavg = as.numeric(formatC(caprisons$cpavg, digits = 2, format = "f"))

# Total failure to abate penalty
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$fta_total = sum(estb$fta_penalty, na.rm = TRUE)
  return(estb)
}))


# Average difference between initial and current penalties

caprisons$ipcpdif = (caprisons$initial_penalty - caprisons$current_penalty)

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$avg_ipcpdif = mean(estb$ipcpdif, na.rm = TRUE)
  return(estb)
}))

# format to the nearest dollar
caprisons$avg_ipcpdif = as.numeric(formatC(caprisons$avg_ipcpdif, digits = 0, format = "f"))

```

## Number of Times Inspected/Violations

```{r}
# Number of times inspected/violations. If they are only in there once, might be bc they were only inspected, no violations. Or they might have been inspected once, again later and have subsequent rows for each violation

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$nr_times_in_data = nrow(estab)
  return(estab)
}))


# Number of total violations (not unique types of violations)
# recode noviol True = 0, False = 1 so we can count the number of violations
caprisons$no_violations = caprisons$noviol
caprisons$no_violations = ifelse(grepl(TRUE, caprisons$no_violations), 0, 1)

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$nr_violations = sum((estab$no_violations))
  return(estab)
}))

```

## Number of Instances

```{r}
# Total number of instances of each violation

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$nr_instances_total = sum(estb$nr_instances, na.rm = TRUE)
  return(estb)
}))

```

## Gravity

```{r}
# A couple ways to think about this. Average gravity - however that does not tell us a sense of the magnitude (number) of the moderate or serious violations.

# Separate out Low, Moderate, and High Gravity
caprisons$high_gravity = ifelse(grepl(10, caprisons$gravity), 1, 0)
caprisons$mod_gravity = ifelse(grepl(5, caprisons$gravity), 1, 0)
caprisons$low_gravity = ifelse(caprisons$gravity == 1, 1, 0)


# Total Serious Gravity
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$high_gravity_sum = sum(estb$high_gravity, na.rm = TRUE)
  return(estb)
}))

# Total Moderate Gravity
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$mod_gravity_sum = sum(estb$mod_gravity, na.rm = TRUE)
  return(estb)
}))

# Total Low Gravity
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$low_gravity_sum = sum(estb$low_gravity, na.rm = TRUE)
  return(estb)
}))


```

## Types of Violations

```{r}
# This variable describes if the violation was other than serious, serious, willful, or a repeat violation. Useful to understand in combo with gravity potentially. Gravity is used to create the penalty amount. While violation type gives us insight into if someone had the potential to become seriously injured. 

# Recode the variables to dummy variables so we can count the number of violation types

caprisons$other_than_serious = ifelse(grepl("O", caprisons$viol_type), 1, 0)
caprisons$repeat_violations = ifelse(grepl("R", caprisons$viol_type), 1, 0)
caprisons$serious_violations = ifelse(grepl("S", caprisons$viol_type), 1, 0)

# Total other than serious
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$other_than_serious_sum = sum(estb$other_than_serious, na.rm = TRUE)
  return(estb)
}))

# Total repeat
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$repeat_violations_sum = sum(estb$repeat_violations)
  return(estb)
}))

# Total serious
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$serious_violations_sum = sum(estb$serious_violations)
  return(estb)
}))


# view(caprisons[, c("estab_name", "viol_type", "repeat_violations", "repeat_violations_sum")])


```

## Inspection Scope

```{r}
# A - complete, B- partial, D - noinspection

caprisons$insp_scope_complete = ifelse(grepl("A", caprisons$insp_scope), 1, 0)
caprisons$insp_scope_partial = ifelse(grepl("B", caprisons$insp_scope), 1, 0)

# Total complete inspections

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_scope_complete_sum = sum(estb$insp_scope_complete, na.rm = TRUE)
  return(estb)
}))

# Total partial inspections

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_scope_partial_sum = sum(estb$insp_scope_partial, na.rm = TRUE)
  return(estb)
}))


```

## Inspection Type

```{r}
# A - Accident, B - Complaint, C - Referral, F - Follow up, G - Unprog rel, H - Planned, J - Unprog other,  M - Fat/Cat
# I think I will combine the unprog rel and other. 

# Recode the variables to dummy variables so we can count the number of inspection types

caprisons$insp_type_accidents = ifelse(grepl("A", caprisons$insp_type), 1, 0)
caprisons$insp_type_complaint = ifelse(grepl("B", caprisons$insp_type), 1, 0)
caprisons$insp_type_referral = ifelse(grepl("C", caprisons$insp_type), 1, 0)
caprisons$insp_type_followup = ifelse(grepl("F", caprisons$insp_type), 1, 0)
caprisons$insp_type_unprogrammed = ifelse(grepl("G", caprisons$insp_type), 1, 0)
caprisons$insp_type_unprogrammed = ifelse(grepl("J", caprisons$insp_type), 1, 0)
caprisons$insp_type_planned = ifelse(grepl("H", caprisons$insp_type), 1, 0)
caprisons$insp_type_fat_cat = ifelse(grepl("M", caprisons$insp_type), 1, 0)


# Total accident
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_accidents_sum = sum(estb$insp_type_accidents, na.rm = TRUE)
  return(estb)
}))

# Total complaint
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_complaint_sum = sum(estb$insp_type_complaint, na.rm = TRUE)
  return(estb)
}))

# Total referrals
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_referral_sum = sum(estb$insp_type_referral, na.rm = TRUE)
  return(estb)
}))

# Total insp_type_followup
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_followup_sum = sum(estb$insp_type_followup, na.rm = TRUE)
  return(estb)
}))

# Total insp_type_unprogrammed
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_unprogrammed_sum = sum(estb$insp_type_unprogrammed, na.rm = TRUE)
  return(estb)
}))


# Total insp_type_planned
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_planned_sum = sum(estb$insp_type_planned, na.rm = TRUE)
  return(estb)
}))

# Total fatality/catastrophes
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$insp_type_fat_cat_sum = sum(estb$insp_type_fat_cat, na.rm = TRUE)
  return(estb)
}))


```

## Number of People Exposed

```{r}
# An important thing to keep in mind is the total could be the same people exposed multiple times for multiple different exposures. So it's total number exposed for all violations.

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$total_exposed = sum(estab$nr_exposed, na.rm = TRUE)
  return(estab)
}))

# view(caprisons[, c("estab_name", "nr_exposed", "total_exposed")])



```

## Average Number of People in Facility

```{r}
# Average Number of People in Facilities
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estb){
  estb$avg_ppl_in_facility = mean(estb$nr_in_estab, na.rm = TRUE)
  return(estb)
}))

  # format to no decimals
   caprisons$avg_ppl_in_facility = as.numeric(formatC(caprisons$avg_ppl_in_facility, digits = 0, format = "f"))

```

## Union Status

```{r}
# Recode
caprisons$union_status_r = caprisons$union_status
caprisons$union_status_r = ifelse(grepl("A", caprisons$union_status_r), "Y", caprisons$union_status_r)
caprisons$union_status_r = ifelse(grepl("B", caprisons$union_status_r), "N", caprisons$union_status_r)
caprisons$union_status_rp = ifelse(grepl("Y", caprisons$union_status_r), 1, 0) # present
caprisons$union_status_rnp = ifelse(grepl("N", caprisons$union_status_r), 1, 0) # not present

# Number of times union ever present during inspection
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$union_present_sum = sum(estab$union_status_rp, na.rm = TRUE)
  return(estab)
}))

# Number of times union not present during inspection
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$union_not_present_sum = sum(estab$union_status_rnp, na.rm = TRUE)
  return(estab)
}))

# union present a majority of the times inspected 
caprisons$union_majority = ifelse((caprisons$union_present_sum > caprisons$union_not_present_sum), 1, 0)


```

# Open and Close Dates

```{r}
# Length of Time Inspection/Violation Open
# Set up
caprisons$open_date = mdy(caprisons$open_date)
caprisons$close_case_date = mdy(caprisons$close_case_date)

caprisons$open_close_interval = caprisons$open_date %--% caprisons$close_case_date
caprisons$open_close_duration = as.duration(caprisons$open_close_interval)
caprisons$open_close_period = as.period(caprisons$open_close_interval)

# Days Open
caprisons$case_days_open = caprisons$open_close_duration / ddays(1)
# Weeks Open
caprisons$case_weeks_open = caprisons$open_close_duration / dweeks(1)
# Months Open
caprisons$case_months_open = caprisons$open_close_duration / dmonths(1)
# Years Open
caprisons$case_years_open = caprisons$open_close_duration / dyears(1)

# Total # Days all Inspections Open
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$case_days_open_sum = sum(estab$case_days_open, na.rm = TRUE)
  return(estab)
}))

# Average # Days all Inspections Open
caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$case_days_open_mean = mean(estab$case_days_open, na.rm = TRUE)
  return(estab)
}))

  # Format to two decimals
  caprisons$case_days_open_mean = as.numeric(formatC(caprisons$case_days_open_mean, digits = 0, format = "f"))

# Check work
# view(caprisons[, c("estab_name", "open_date", "close_case_date", "case_days_open", "case_days_open_sum", "case_days_open_mean")])

```

# Data Cleaning Steps

## Hazardous Substances, Standards, and Activity Nrs

```{r}
# Recode the hazardous substances

caprisons$hazsub1_recode = factor(caprisons$hazsub1,
                                  levels = c("1340", "1430", "8120", 
                                             "8870", "8880", "B418", "S227"),
                                  labels = c("1340 - Gasoline", 
                                             "1430 - Hydrogen Chloride",
                                             "8120 - Fungi",
                                             "8870 - Electrical Shock", 
                                             "8880 - All Other Physical Hazards", 
                                             "B418 - Bacteria", 
                                             "S227 - Sodium Hypochlorite"))

caprisons$hazsub2_recode = factor(caprisons$hazsub2,
                                  levels = c("8880", "9523"),
                                  labels = c("8880 - All Other Physical Hazards",
                                             "9523 - Black Mold"))



caprisons$hazsub3_recode = factor(caprisons$hazsub3,
                                  levels = c("8870"),
                                  labels = c("8870 - Electrical Shock"))
# reassign as class character
caprisons$hazsub1_recode = as.character(caprisons$hazsub1_recode)
caprisons$hazsub2_recode = as.character(caprisons$hazsub2_recode)
caprisons$hazsub3_recode = as.character(caprisons$hazsub3_recode)


# Clean standards column

# trim white space
caprisons$standardc = trimws(caprisons$standard)
# remove ()
caprisons$standardc = gsub("\\(", "", caprisons$standardc)
caprisons$standardc = gsub("\\)", "", caprisons$standardc)
# remove spaces
caprisons$standardc = gsub(" ", "", caprisons$standardc)
# split column at letter
caprisons$standardnumber = caprisons$standardc
caprisons$standardnumber = gsub("[A-Za-z].*", "", caprisons$standardnumber)


# Change class of activity nr
caprisons$activity_nr = as.character(caprisons$activity_nr)

```

## Zipcodes

```{r}
caprisons$site_zip = as.character(caprisons$site_zip)

```

## Owner Type

```{r}
# Some of the facilities are listing different owner types. I review those facilities and assign them an owner type based on what I can determine from googling. 

caprisons[which(caprisons$estab_name == "BARRY J NIDORF JUVENILE HALL"), "owner_type"] = "B"
caprisons[which(caprisons$estab_name == "CA DEPT OF STATE HOSPITALS"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "CALIFORNIA INSTITUTION FOR MEN (CIM)"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "CALIFORNIA STATE PRISON-LOS ANGELES COUNTY (LAC)"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "CALIPATRIA STATE PRISON"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "CHUCKAWALLA VALLEY STATE PRISON (CVSP)"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "DEUEL VOCATIONAL INSTITUTION (DVI)"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "IRONWOOD STATE PRISON (ISP)"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "N.A. CHADERJIAN YOUTH CORRECTIONAL FACILITY"), "owner_type"] = "C"
caprisons[which(caprisons$estab_name == "PITCHESS DETENTION CENTER"), "owner_type"] = "B"
caprisons[which(caprisons$estab_name == "VENTURA YOUTH CORRECTIONAL FACILITY"), "owner_type"] = "C"

# Recode 
caprisons$owner_typef = caprisons$owner_type
caprisons$owner_typef = factor(caprisons$owner_typef,
                                  levels = c("A", "B", "C", "D"),
                                  labels = c("PRIVATE",
                                             "LOCAL GOVERNMENT",
                                             "STATE GOVERNMENT",
                                             "FEDERAL GOVERNMENT"))



```

# Reshape the Data to Wide Format

```{r}
# reshape the dataset to wide and keep only toxicity variables

caprisons = do.call(rbind, by(caprisons, caprisons$estab_name, FUN = function(estab){
  estab$estab_count = 1:nrow(estab)
  return(estab)
}))

keepcolumns <- c("estab_name",  "site_address", "site_city", "site_state", "site_zip","estab_count", "nr_times_in_data", "nr_violations", "nr_instances_total", "total_exposed", "avg_ppl_in_facility", "low_gravity_sum", "mod_gravity_sum", "high_gravity_sum","other_than_serious_sum", "repeat_violations_sum", "serious_violations_sum", "cptotal", "iptotal", "ipavg", "cpavg", "avg_ipcpdif", "fta_total", "hazsub1_recode", "hazsub2_recode", "hazsub3_recode", "standardnumber", "owner_typef", "union_present_sum", "union_not_present_sum", "union_majority", "case_days_open_sum", "case_days_open_mean", "insp_type_accidents_sum", "insp_type_complaint_sum", "insp_type_referral_sum", "insp_type_followup_sum", "insp_type_unprogrammed_sum", "insp_type_planned_sum", "insp_type_fat_cat", "insp_scope_complete_sum", "insp_scope_partial_sum", "activity_nr")


caprisonsk = caprisons[keepcolumns]

caprisons_wide = reshape(caprisonsk, idvar = "estab_name", timevar = "estab_count", direction = "wide")

row.names(caprisons_wide) <- 1:nrow(caprisons_wide)


```

# Compile list of all Hazards and Standards Reported

## Hazardous Substances

```{r}
# unique first then paste

all_haz_string = apply(caprisons_wide, 1, FUN = function(vrow) {
  
  # get all haz columns and get the unique hazardous substances
  haz1 = vrow[names(vrow)[grepl("hazsub1_recode\\.\\d", names(vrow), perl = TRUE)]]
  haz1 = unique(haz1)
  
  haz2 = vrow[names(vrow)[grepl("hazsub2_recode\\.\\d", names(vrow), perl = TRUE)]]
  haz2 = unique(haz2)
  
  haz3 = vrow[names(vrow)[grepl("hazsub3_recode\\.\\d", names(vrow), perl = TRUE)]]
  haz3 = unique(haz3)
  
  # make sure unique across all three hazsub variables
  all_haz = unique(c(haz1, haz2, haz3))
  
  # combine into single string
  haz_string = paste0(all_haz[!is.na(all_haz)], collapse = ", ")
  
  # turn string NAs into real NAs
  if(haz_string == ""){haz_string = NA}
  
  # return
  return(haz_string)
})

# assign back to dataframe
caprisons_wide$haz_string = all_haz_string


```

## Standards

```{r}

# unique first then paste

all_standards_string = apply(caprisons_wide, 1, FUN = function(vrow) {
  
  # get all haz columns
  stndrds = vrow[names(vrow)[grepl("standardnumber\\.\\d", names(vrow), perl = TRUE)]]
  stndrds = unique(stndrds)
  
  # combine into single string
  stndrds_string = paste0(stndrds[!is.na(stndrds)], collapse = ", ")
  
  # turn string NAs into real NAs
  if(stndrds_string == ""){stndrds_string = NA}
  
  # return
  return(stndrds_string)
})

# assign back to dataframe
caprisons_wide$standards_string = all_standards_string




```

# Create list of activity nrs and count of times investigated

```{r}

# unique first then paste

all_activity_string = apply(caprisons_wide, 1, FUN = function(vrow) {
  
  # get all haz columns
  actv = vrow[names(vrow)[grepl("activity_nr\\.\\d", names(vrow), perl = TRUE)]]
  actv = unique(actv)
  
  # combine into single string
  actv_string = paste0(actv[!is.na(actv)], collapse = ", ")
  
  # turn string NAs into real NAs
  if(actv_string == ""){actv_string = NA}
  
  # return
  return(actv_string)
})

# assign back to dataframe
caprisons_wide$actv_string = all_activity_string


# counts the number of activity_nrs to get count of total times investigated
invest_count = caprisons %>% group_by(estab_name) %>% summarise(count = n_distinct(activity_nr))
caprisons_wide$nr_investigations = invest_count$count

```

# Subset the Wide Data to Only Show Desired Columns

```{r}
# For full wide data
#caprisons_dt = caprisons_wide %>% select((c("estab_name","site_address.1", "site_city.1", "site_state.1", "site_zip.1", "owner_typef.1", "nr_times_in_data.1", "nr_violations.1", "nr_instances_total.1", "case_days_open_sum.1", "case_days_open_mean.1", "union_present_sum.1", "union_not_present_sum.1", "union_majority.1", "insp_type_accidents_sum.1", "insp_type_complaint_sum.1", "insp_type_referral_sum.1", "insp_type_followup_sum.1", "insp_type_unprogrammed_sum.1", "insp_type_planned_sum.1", "insp_type_fat_cat.1", "insp_scope_complete_sum.1", "insp_scope_partial_sum.1", "low_gravity_sum.1", "mod_gravity_sum.1", "high_gravity_sum.1", "other_than_serious_sum.1", "repeat_violations_sum.1", "serious_violations_sum.1", "avg_ppl_in_facility.1", "total_exposed.1", "iptotal.1", "cptotal.1", "ipavg.1", "cpavg.1",  "fta_total.1", "haz_string", "standards_string", "nr_investigations")))

# For Table

caprisons_dt = caprisons_wide %>% select((c("estab_name","site_address.1", "site_city.1", "site_state.1", "site_zip.1", "owner_typef.1", "actv_string", "standards_string", "haz_string", "nr_investigations", "nr_violations.1", "nr_instances_total.1", "case_days_open_sum.1", "case_days_open_mean.1", "low_gravity_sum.1", "mod_gravity_sum.1", "high_gravity_sum.1", "other_than_serious_sum.1", "repeat_violations_sum.1", "serious_violations_sum.1", "avg_ppl_in_facility.1", "total_exposed.1", "iptotal.1", "cptotal.1", "ipavg.1", "cpavg.1",  "fta_total.1", "union_present_sum.1", "union_not_present_sum.1", "insp_type_accidents_sum.1", "insp_type_complaint_sum.1", "insp_type_referral_sum.1", "insp_type_followup_sum.1", "insp_type_unprogrammed_sum.1", "insp_type_planned_sum.1", "insp_type_fat_cat.1",  "insp_scope_complete_sum.1", "insp_scope_partial_sum.1", "avg_ipcpdif.1")))

```

## Set up the DT Table

```{r}
dt = datatable(caprisons_dt,
          class = "display",
          rownames = FALSE,
          filter = "top",
          colnames = c("Facility", "Address", "City", "State", "Zipcode", "Facility Type",
                       "Inspection Number(s)",
                       "All Standards Cited",
                       "Any Reported Hazardous Substances",
                       "Number of Inspections", "Number of Violations",
                       "Total Number of Citation Instances", 
                       "Total Days all Inspections Open", 
                       "Average Number of Days all Inspections Open",
                       "Total Low Gravity Violation",
                       "Total Moderate Gravity Violations", 
                       "Total High Gravity Violations", 
                       "Total 'Other than Serious' Violations", 
                       "Total Repeat Violations", "Total Serious Violations", 
                       "Average # of People in Facility", "Total Ever Exposed", 
                       "Total Initial Penalty", "Total Current Penalty", 
                       "Average Initial Penalty", "Average Current Penalty" , 
                       "Failure to Abate Penalty",
                       "Times Union Present at Inspection",
                       "Times Union Not Present at Inspection",
                       "Total Inspections Triggered by Accident", 
                       "Total Inspections Triggered by Complaint", 
                       "Total Inspections Triggered by Referral", 
                       "Total Inspections From a Follow Up", 
                       "Total Unprogrammed Inspections", 
                       "Total Planned Inspections", 
                       "Total Inspections Triggered by Fatality or Catastrophe",
                       "Total Complete Inspections",
                       "Total Partial Inspections"),
            caption = 
            htmltools::tags$caption("\n OSHA Inspections and Violations Data for California Carceral Facilities 2009-2019", 
            style = "color:black; caption-side: top; text-align:left; font-size: 125%; font-weight = bolder"),
          options = 
          list(
          pageLength = 25,
          lengthMenu = c(25, 50, 75, 100, 150),
          initComplete = JS("function(settings, json) {",
       "$(this.api().table().header()).css({'background-color': '#fff', 'color': '#000','font-size': '100%'});",
    "}")
)) %>% 
  formatStyle(
    "iptotal.1",
    background = styleColorBar(caprisons_dt$iptotal.1, color = 'steelblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) %>%
   formatStyle(
    "cptotal.1",
    background = styleColorBar(caprisons_dt$cptotal.1, color = 'lightblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "ipavg.1",
    background = styleColorBar(caprisons_dt$ipavg.1, color = 'steelblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "cpavg.1",
    background = styleColorBar(caprisons_dt$cpavg.1, color = 'lightblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
   formatStyle(
    "avg_ppl_in_facility.1",
    background = styleColorBar(caprisons_dt$avg_ppl_in_facility.1, color = '#2ca25f'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
    formatStyle(
    "total_exposed.1",
    background = styleColorBar(caprisons_dt$total_exposed.1, color = '#99d8c9'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "nr_investigations",
    background = styleColorBar(caprisons_dt$nr_investigations, color = '#756bb1'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
   formatStyle(
    "nr_violations.1",
    background = styleColorBar(caprisons_dt$nr_violations.1, color = '#bcbddc'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
   formatStyle(
    "nr_instances_total.1",
    background = styleColorBar(caprisons_dt$nr_instances_total.1, color = '#efedf5'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "low_gravity_sum.1",
    background = styleColorBar(caprisons_dt$low_gravity_sum.1, color = '#fee8c8'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "mod_gravity_sum.1",
    background = styleColorBar(caprisons_dt$mod_gravity_sum.1, color = '#fdbb84'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "high_gravity_sum.1",
    background = styleColorBar(caprisons_dt$high_gravity_sum.1, color = '#e34a33'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "serious_violations_sum.1",
    background = styleColorBar(caprisons_dt$serious_violations_sum.1, color = '#de2d26'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "repeat_violations_sum.1",
    background = styleColorBar(caprisons_dt$repeat_violations_sum.1, color = '#fc9272'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>%
  formatStyle(
    "other_than_serious_sum.1",
    background = styleColorBar(caprisons_dt$other_than_serious_sum.1, color = '#fee0d2'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
    formatStyle(
    "case_days_open_sum.1",
    background = styleColorBar(caprisons_dt$case_days_open_sum.1, color = '#5E8B7E'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
     formatStyle(
    "case_days_open_mean.1",
    background = styleColorBar(caprisons_dt$case_days_open_mean.1, color = '#A7C4BC'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
    formatStyle(
   "union_present_sum.1",
    background = styleColorBar(caprisons_dt$union_present_sum.1, color = '#f25f5a'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "union_not_present_sum.1",
    background = styleColorBar(caprisons_dt$union_not_present_sum.1, color = '#f58682'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
    formatStyle(
   "insp_type_accidents_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_accidents_sum.1, color = '#3C5186'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_type_complaint_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_complaint_sum.1, color = '#9B72AA'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_type_referral_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_referral_sum.1, color = '#5D8233'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
     formatStyle(
   "insp_type_followup_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_followup_sum.1, color = '#FFF5DE'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_type_unprogrammed_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_unprogrammed_sum.1, color = '#93D9A3'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_type_planned_sum.1",
    background = styleColorBar(caprisons_dt$insp_type_planned_sum.1, color = '#ED8E7C'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_type_fat_cat.1",
    background = styleColorBar(caprisons_dt$insp_type_fat_cat.1, color = '#A03C78'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
  formatStyle(
   "fta_total.1",
    background = styleColorBar(caprisons_dt$fta_total.1, color = '#53B8BB'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_scope_complete_sum.1",
    background = styleColorBar(caprisons_dt$insp_scope_complete_sum.1, color = '#F7A440'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
   formatStyle(
   "insp_scope_partial_sum.1",
    background = styleColorBar(caprisons_dt$insp_scope_partial_sum.1, color = '#F6DCBF'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
   ) %>% 
  formatStyle(columns = c("standards_string", "haz_string", "actv_string"), fontSize = '75%') %>%
  formatStyle("standards_string","white-space"="normal") %>% 
  formatCurrency(columns = c("iptotal.1", "cptotal.1", "ipavg.1", "cpavg.1", "fta_total.1"), currency = "$")
 


```

# Carceral Facilities in the OSHA Data from 2009 to 2019 and Their Record for Hazards and Toxicity

```{r}
dt

```

# Export Wide Data

```{r}

write.csv(caprisons_dt, file = "CA_prison_insp_viol_2010_op_fac_wide.csv", row.names = FALSE)
```
